Extended Data

### 2021-11-16 doc/manuscript/Dow_manuscript_Nature.Rmd
edit first par of discussion

Together, our results demonstrate that warmer spring temperatures in the temperate deciduous forests of eastern North America advance the phenology of tree stem growth but have little effect on annual woody productivity (Figs. `r schematic_summary`- `r tree_rings`). 
The observed phenological advance in the start of stem growth under warmer springs parallels phenological advances observed for canopy foliage (Fig. `r DOY_timing`a-b)[@keenan_net_2014;@finzi_carbon_2020;@stuble_plant_2021] and _NEE_ [@keenan_net_2014; @finzi_carbon_2020].
However, inconsistent with the concept that an earlier start to growth would increase annual woody productivity, we demonstrate that warmer springs hasten the cessation of stem expansion, thus having negligible effect on total annual growth for most species and locations (Fig. `r tree_rings`). 
Our observations suggest that the cessation of rapid stem expansion, which occurs mid-summer near the time of peak canopy greenness (`r EDF_growth_curves_all`)[@finzi_carbon_2020], is likely driven by cues other than carbon limitation, such as daylength or sink limitation, which also play an important role in autumn leaf senescence[@zani_increased_2020;@keenan_timing_2015;@fu_variation_2014].
Rather than lengthening the growing season, the primary effect of warming temperature on annual growth is a negative correlation with $T_{max}$ and associated drought stress during the peak growing season[@dorangeville_drought_2018], offsetting any benefit of an earlier onset of growth in the spring[@zhang_drought_2021].
Warm springs may also augment summer drought stress in some times and places, effectively canceling out any positive effects of an extended growing period[@buermann_widespread_2018]; however, spring temperatures and summer Standardized Precipitation Evapotranspiration Index[@vicente-serrano_multiscalar_2010] were uncorrelated within our dendrometer band analysis.
The effect of warm springs specifically in years with higher levels of precipitation during the peak growing season is not elucidated from this study, and should be examined in the future to clarify this relationship.

>>>>

Together, our results demonstrate that warmer spring temperatures in the temperate deciduous forests of eastern North America advance the phenology of tree stem growth but have little effect on annual woody productivity (Figs. `r schematic_summary`- `r tree_rings`). 
The observed phenological advance in the start of stem growth under warmer springs parallels phenological advances observed for canopy foliage (Fig. `r DOY_timing`a-b)[@keenan_net_2014;@finzi_carbon_2020;@stuble_plant_2021] and _NEE_[@keenan_net_2014; @finzi_carbon_2020].
However, inconsistent with the concept that an earlier start to growth would increase annual woody productivity, we demonstrate that warmer springs hasten the cessation of stem expansion and thereby have negligible effect on total annual growth for most species and locations (Fig. `r tree_rings`). 
Our observations suggest that the cessation of rapid stem expansion, which occurs mid-summer near the time of peak canopy greenness (`r EDF_growth_curves_all`)[@finzi_carbon_2020], is likely driven by cues other than carbon limitation, such as daylength or sink limitation, which also play an important role in autumn leaf senescence[@zani_increased_2020;@keenan_timing_2015;@fu_variation_2014].
Our tree-ring analysis (Fig. `r tree_rings`) demonstrates that the primary effect of warming temperatures on annual tree growth is not an augmentation through an earlier start to growth, but rather a reduction associated to drought stress during the peak growing season[@dorangeville_drought_2018].
Warm springs may also amplify summer drought stress in some times and places, effectively canceling out any positive effects of an extended growing period[@buermann_widespread_2018;@zhang_drought_2021]; however, spring temperatures and summer Standardized Precipitation Evapotranspiration Index[@vicente-serrano_multiscalar_2010] were uncorrelated within our dendrometer band analysis, implying that the effects of warm spring temperatures on growth phenology elucidated here (Fig. `r schematic_summary`) were not attributable to summer drought.
